Full Article Review

### 2024-10-21 doc/reviews/luke.qmd
docs: read through more papers

@@

@@

@@

-   While the project's output is a data sharing platform, the process
    to get there required a high level of collaboration.
-   Seems to use GitHub to at least build the website.
-   Building up an open collaboration project takes years of continued
    work and effort.
-   Took a multi-staged approach to establishing the group/community.
    -   At each stage, it was slowly about introducing the concepts to a
        broader and broader audience and getting feedback all along the
        way.
-   Development follows principles of agile software development by
    getting requirements from users, designing and developing those
    requirements, seeking feedback from users on what was developed, and
    testing the developed features.
-   Use of the iterative development helped them identify and respond to
    issues that came up. This is something that could be something that
    non-software projects could use to help with collaboration and
    maintaining momentum and motivation.

@@

-   Mostly an example of what they did and why
-   Faced with limited resources and small sample sizes, but
    wanting/needing more to make results generalizable, they put
    together this collaboration project.
-   Outlines benefits of this collaboration, such as reducing the
    efforts of individual researchers by distributing workload,
    resources, and knowledge sharing/creation/management.
-   Challenges to this collaboration are less papers, lower ability to
    cover more topics, less freedom of control and direction at the
    individual researcher level (since they are part of a larger group
    and need to have a consensus on a topic or standards of practice),
    higher level of rules and processes for dealing with researchers who
    try to take advantage of the community for their own benefit, as
    well as how to decide on authorship/contributorship for scientific
    publications.
    -   A big challenge to these collaborations is sources of funding.
        Who pays for whose salary? How are funds distributed?
-   Argues that any large-scale collaboration needs to use open
    practices by default.
-   Something to consider when building large-scale collaboration is its
    continuation. If its started by one or a few people and with
    dedicated funding for it, what happens if that person or people
    leave or move on to other things? Or when the funding ends? These
    need to be considered at the start.
-   The ManyPrimates use GitHub, at least to host their website with
    some general documentation and guidelines, as well as some analysis
    code and data for some papers they've published.
-   The ManyPrimates project also set some basic standards for project
    file and folder structure, for pre-registering the study, etc.

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

-   Systematic review
-   Large-scale collaboration
-   Large-scale infrastructure

@@

-   Review
-   Example workflow

@@

-   This paper wasn't the most relevant, as it didn't go into as much
    detail as we might like, even though it was describing exactly what
    we are looking for.

-   Research needs a certain sample size to be able to answer certain
    questions. But primate research is difficult to get enough sample
    size for any given project. This need drove the push for more
    openness and sharing.

-   Centralized platform for primate researchers and for sharing data.

-   Use GitHub to share things.

-   Projects should preregister the plans, and all data, code, and
    materials are uploaded to GitHub.

-   Tools:

    -   GitHub
    -   Google Docs
    -   Slack
    -   OSF

-   Use their website to disseminate news and findings

-   Limitations:

    -   Logistical. Impossible to use the exact same methods across all
        studies.
    -   Also technical and knowledge, not enough funding to hire
        relevant expertise and technical resources.

@@

-   While the project's output is a data sharing platform, the process
    to get there required a high level of collaboration.

-   Seems to use GitHub to at least build the website.

-   Building up an open collaboration project takes years of continued
    work and effort.

-   Took a multi-staged approach to establishing the group/community.

    -   At each stage, it was slowly about introducing the concepts to a
        broader and broader audience and getting feedback all along the
        way.

-   Development follows principles of agile software development by
    getting requirements from users, designing and developing those
    requirements, seeking feedback from users on what was developed, and
    testing the developed features.

-   Use of the iterative development helped them identify and respond to
    issues that came up. This is something that could be something that
    non-software projects could use to help with collaboration and
    maintaining momentum and motivation.

@@

-   Mostly an example of what they did and why

-   Faced with limited resources and small sample sizes, but
    wanting/needing more to make results generalizable, they put
    together this collaboration project.

-   Outlines benefits of this collaboration, such as reducing the
    efforts of individual researchers by distributing workload,
    resources, and knowledge sharing/creation/management.

-   Challenges to this collaboration are less papers, lower ability to
    cover more topics, less freedom of control and direction at the
    individual researcher level (since they are part of a larger group
    and need to have a consensus on a topic or standards of practice),
    higher level of rules and processes for dealing with researchers who
    try to take advantage of the community for their own benefit, as
    well as how to decide on authorship/contributorship for scientific
    publications.

    -   A big challenge to these collaborations is sources of funding.
        Who pays for whose salary? How are funds distributed?

-   Argues that any large-scale collaboration needs to use open
    practices by default.

-   Something to consider when building large-scale collaboration is its
    continuation. If its started by one or a few people and with
    dedicated funding for it, what happens if that person or people
    leave or move on to other things? Or when the funding ends? These
    need to be considered at the start.

-   The ManyPrimates use GitHub, at least to host their website with
    some general documentation and guidelines, as well as some analysis
    code and data for some papers they've published.

-   The ManyPrimates project also set some basic standards for project
    file and folder structure, for pre-registering the study, etc.

@@

-   Example
-   Workflows
-   Barriers
-   Lessons learned
-   Tools

@@

-   Example of a workflow
-   Guide

@@

-   This was a very good paper and very relevant and detailed.

-   More inclusive term might be "open scholarship"

-   Applying openness can be very challenging when it involves data on
    children, but also that data on children is difficult to collect and
    benefits substantially from researchers using open practices,
    especially open data.

-   The WomCogDev lab is a fairly small group (\~13 people), almost all
    of whom are students (e.g. Masters, undergraduates, PhD students).

-   General steps in their workflow:

    1.  Design project
    2.  Pre-register project or Registered Report
    3.  Create experiment
    4.  Recruit participants and collect data
    5.  Analyze data
    6.  Disseminate results

-   Focus on 1, 4, and 5 since they say those are less discussed in the
    literature.

-   For 1, involves writing something up by the lead researcher, getting
    feedback from all lab, writing detailed plan, upload to OSF, decide
    on authorship.

    -   Making hypothesis and thinking of what all possible outcomes for
        results and what that means in the context of the hypothesis
    -   A Data Management Plan is also created at this stage that
        includes details about how to be FAIR

-   For 2, they start off with trying to do a Registered Report, but if
    it gets rejected, they will take the reviewer comments, revise and
    make edits, and then upload to OSF as a pre-registered protocol.

-   For 3, they do not have a common programming language for creating
    experiments but do prioritize open source over proprietary.

-   For 4, something they do to help with reproducibility is to use
    "data partners", which is someone who is in the lab but not directly
    involved in the specific project. Their task is to re-analyse the
    data to see if they come to the same conclusions as the lead author.
    Generally, the data partner is a PhD student or more senior.

-   Tools used:

    -   Slack for communication
    -   Trello board
    -   OSF, which has components made for literature, documents,
        methods, raw data, experiment details, analysis and results,
        manuscript, and communication (e.g. for outreach). Made private
        at first.
    -   R or JASP for analysis
    -   bioRxiv/PsyArXiv for preprints

-   Philosophy:

    -   More work done upfront, the less work done later.
    -   Some openness is better than no openness.
    -   Openness is on a spectrum.
    -   Always improving and can't always do everything at once.

-   Other workflows notes:

    -   Each weekly lab meeting, people start with "an interesting error
        I found..." to encourage self-correction and open and honest
        dialogue and discussions
    -   They have a lab manual which describes these processes.
    -   Not all projects follow this workflow, nor are they open.
    -   Try to build the values into the work culture

-   Advantages of workflow:

    -   Uncertainty is reduced
    -   Reproducibility is inherent to the design and practice of the
        workflow

-   Disadvantages of workflow:

    -   Can sometimes be quite time-consuming
    -   Can be difficult to get everyone on board and to follow it
        fairly closely
    -   Steep learning curve
    -   Overreliance on OSF, which has a cap on amount of data to put
        into it

@@

-   Tips
-   Guidelines
-   Review
-   Strategies
-   Overview

@@

-   Community-building
-   Tips
-   Guide

@@

-   Relevant paper as an overview

-   Open science resources tend to be more theoretical then practical
    ("do this" but not "this is how you do this")

-   Strategy: Start small and slowly incorporate open science practices.

    -   For instance, start with uploading a preprint, then
        pre-register, then share notebooks/methods, then analysis code,
        etc.

-   Strategy: Become familiar with national/institutional open science
    policies

-   Strategy: Analyze and share successful cases of implementing open
    science practices

    -   By sharing, you bring awareness to others about it, which can
        over time pressure and push for institutional/organizational
        change.

-   Strategy: Include open science practices in your teaching

    -   Getting students to be more familiar with and comfort on open
        science practices, in course settings and assignments, will mean
        they will be more comfortable applying it to real projects.

-   Strategy: Embrace and use open educational resources

    -   Use them, create them, publish them

-   Strategy: Collaborate with others using open tools

    -   Example: Write an article using R Markdown or Quarto. Use open
        formats for data. Use GitHub to share and collaborate on
        projects.

-   Strategy: Develop networks of open collaboration

    -   Build networks of working with others in an open way.
    -   Develop standards, frameworks, and templates to follow within
        this network.

-   Strategy: Voice your opinions

    -   Write things in support of these initiatives
    -   Outline values and principles to follow

-   Strategy: Rethink and promote changes in assessment of scholarship

    -   Push for and advocate for alternatives to assessments outside of
        the typical impact factor and h-index

-   Strategy: Create opportunities for people to specialize in open
    science

-   Tools:

    -   OSF, GitHub, Zenodo, R, Quarto, R Markdown, Jupyter Notebooks

@@

-   Best practices
-   Lessons learned
-   Review

@@

-   Example of workflow
-   Guide

@@

-   Very general outline of what they've done, without very clear
    details on how they did it

-   Best practices: Pre-registration, FAIR data, reproducible imaging,
    open science of workflows, methods, analysis code

-   Strategies:

    -   Include a long review period to identify best practices and
        tools before implementing any changes
    -   Implement version control using Git and GitHub, so that when
        students move on, the knowledge isn't lost
    -   Transition from Matlab to Python (or R)
    -   For new projects, start from scratch using best practices
    -   For older projects, update to best practices as much as possible
        given time and effort
    -   All finished papers need to be uploaded to preprint archives
    -   Convert data to BIDS format (very difficult procedure)
    -   Converted in-house pipeline to standard fMRIPrep, which is a
        standard imaging pipeline
    -   Upload data to OpenNeuro
    -   Constructed a common data dictionary for data collected (very
        difficult)

-   Tools:

    -   Git, GitHub, Python
    -   psyRxiv, bioRxiv, arXiv
    -   BIDS data format as well as BIDS-Apps
    -   fMRIPrep and Singularity
    -   Deposit data on OpenNeuro

-   Benefits:

    -   Uploading to open records/archives shows institutional work and
        knowledge put into the work
    -   Reduced time to completion of many tasks
    -   Reduced costs
    -   Minimal loss of knowledge when students leave, so knowledge can
        more easily be built up on more
    -   Reducing errors from processing, analyzing, and data collection
    -   Increased cross-group collaboration and knowledge sharing
    -   Uploading preprints, using open source programming languages,
        depositing data to public repos, and using containerized
        reproducible pipelines all had low cost-high benefit ratios.
